---
title: "Data Visualization Using ggplot"
output: html_document
---

Let's work on the Lahman Dataset with SQL and ggplot()
First let's add the packages.
```{r}
library("Lahman", lib.loc="~/R/win-library/3.4")
library("sqldf", lib.loc="~/R/win-library/3.4")
library("ggplot2", lib.loc="~/R/win-library/3.4")
```

Next, let's SQL some data into a query.
```{r}
#Data visulization scatter plots
#extracting Data 
query <- "SELECT playerID, sum(HR) AS Career_HR, sum(SO) AS Career_SO
FROM Batting 
GROUP BY playerID 
HAVING sum(HR)>=400"
sqldf(query)


```

To display the data, you then need to create a dataframe.
```{r}

plotdf <- data.frame(sqldf(query))
```

Now we can make some graphs.
First we will start with a scatter plot with the ggpolt() package.
There are three key elements to ggplot: Data, Aestetics and Geometery.
```{r}
plot1 <- ggplot(data=plotdf, aes(x=Career_SO, y=Career_HR, label=playerID)) + 
  geom_point()+
  geom_text(aes(label=playerID),hjust=0, vjust=0)+
  ggtitle("Career Homeruns VS Career Strikeouts for Great Hitters")+
  xlab("Career Strikeouts")+
  ylab("Career Home Runs")
plot1
```

We can also make a histogram using player weight 
```{r}
#Histogram ggplot

#Extracting Data
query2 <- "SELECT weight FROM Master"
plotdf_2<-sqldf(query2)

#Data Viz

plot2 <- ggplot()+
  geom_histogram(data=plotdf_2, aes(x=weight), color="red", fill="white")+
  ggtitle("Ball Player Bodyweight Distribution")
plot2
```

We can also combine aspects of ggplot() in order to make a time series graph
```{r}
#Time Series Plot for the Great Bambino 

#Extracting Data
query3 <- "select yearID, HR from Batting
where playerID='ruthba01'"

plotdf_3<-sqldf(query3)

#Data Viz 

plot3 <- ggplot()+
  geom_line(data=plotdf_3, aes(x=yearID, y=HR),color='blue')+
  geom_point(data=plotdf_3, aes(x=yearID, y=HR),color='red')+
  ggtitle("The Sultan of Swat's Homerun Blasts Through the Years")+
  xlab("Year")+
  ylab("Homerun Blasts")
plot3
```

We can also make some pretty niffty bar plots using the same process
```{r}
#barplot
#extracting the data
colnames(Teams)

query4 <- 'select name, HR from Teams
where yearID=1980
order by HR'
plotdf_4 <- sqldf(query4)

#making the barplot 

plotdf_4$name<-factor(plotdf_4$name, levels=plotdf_4$name)


plot4 <- ggplot(data=plotdf_4, aes(x=name, y=HR, fill=HR))+
  geom_bar(stat='identity')+
  xlab('Homeruns')+
  ylab('Team Names')+
  ggtitle('1980 Team Homerun Distribution')+
  coord_flip()
  
plot4
```

